---
title: "Reproducible Research Peer Assignment 1"
author: "Kshitij Makwana"
---

## Loading and Pre-processing the data 

```{r loading, echo=TRUE}
library(chron)
library(dplyr)
library(ggplot2)
library(reshape2)
activity<-read.csv("activity.csv")
```

## What is mean total number of steps taken each day?

```{r mean_steps, echo=TRUE}
tot_steps1<-tapply(activity$steps,as.factor(activity$date),sum,na.rm=TRUE)
plot(tot_steps1,type = "h", xaxt="n",xlab = "Date",ylab = "Steps")
axis(1,at=c(1:dim(tot_steps1)),labels = names(tot_steps1))
mean_steps1<-mean(tot_steps1)
median_steps1<-median(tot_steps1)
print(paste("Mean steps in a day are",round(mean_steps1)))
print(paste("Median of total number steps is ",median_steps1))
```

## What is the average daily activity pattern?

```{r mean_interval_activity, echo=TRUE}
mean_interval<-tapply(activity$steps,activity$interval,mean,na.rm=TRUE)
plot(mean_interval,xaxt="n",xlab = "Time Interval", ylab = "Mean Steps",type = "l")
axis(1,at=c(1:length(mean_interval)),labels = names(mean_interval))
print(paste("Maximum average steps is for the time interval",names(which.max(mean_interval)),"s"))
```

## Imputing Missing Values

```{r imputing_na, echo=TRUE}
activity_copy<-activity
print(paste("Total number of missing values:",sum(is.na(activity_copy))))
mean_daily<-tapply(activity_copy$steps,as.factor(activity_copy$date),
                   mean,na.rm=TRUE)
for(i in 1:length(activity_copy$steps)){
  if(is.na(activity_copy$steps[i])){
    if(!(is.nan(mean_daily[activity_copy$date[i]]))){
      activity_copy$steps[i]<-mean_daily[activity_copy$date[i]]
    }
    else{
      activity_copy$steps[i]<-0
    }
  }
}

tot_steps<-tapply(activity_copy$steps,as.factor(activity_copy$date),sum)
plot(tot_steps,type = "h", xaxt="n",xlab = "Date",ylab = "Steps")
axis(1,at=c(1:dim(tot_steps)),labels = names(tot_steps))

mean_steps<-mean(tot_steps)
median_steps<-median(tot_steps)
print(paste("Mean steps in a day are",round(mean_steps)))
print(paste("Median of total number steps is ",median_steps))
```

There is no considerable difference from the estimates from the first part of the assignment. This is mostly because I have replaced the NAs with the mean of the steps in the day.

## Are there differences in activity patterns between weekdays and weekends?

```{r weekdays_and_weekends, echo=TRUE}
for(i in 1:length(activity_copy$date)){
  if(is.weekend(activity_copy$date[i])){
    activity_copy$day[i]<-"weekend"
  }
  else{
    activity_copy$day[i]<-"weekday"
  }
}
activity_copy$day<-as.factor(activity_copy$day)

# weekday<-filter(activity_copy,day=="weekday")
# weekend<-filter(activity_copy,day=="weekend")

data<-dcast(activity_copy,interval~day, fun.aggregate = mean, value.var = "steps")
data<-melt(data,id="interval",variable.name = "day", value.name = "steps")
g<-ggplot(data,aes(interval,steps))
g+geom_line()+facet_grid(day~.)
```

It looks like on a weekday, many steps are taken in the early intervals which might be due to an early running session or rushing to work. There is not much activity for the rest of the day.  
On a weekend, the steps are distributed almost equally throughout he day.